---
title: "Calculating Percent of Calories from Carboydrates"
author: "Bob Horton"
date: "October 20, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Converting from foods to nutrients

To generate a list of nutrients from a set of foods, we add up all the nutrients in all the foods. This is basically a dot product, or matrix multiplication problem, which maps well to SQL. Here we demonstrate the logic using a SQL database; for large dateasets the same logic (and almost exactly the same SQL code) can be applied in Hive on Hadoop.


```{r load_data}
food_nutrients <- read.csv("food_nutrients.csv", header=FALSE, col.names=c("item", "nutrient", "quantity"))
grocery <- read.csv("grocery_sample.csv", header=FALSE, col.names=c("diet_id", "item", "quantity", "units"))

library(sqldf)

sql <- "with
nut as (
	select diet_id, nutrient, sum(fn.quantity * g.quantity/100) as quantity 
		from grocery g join food_nutrients fn 
		on g.item = fn.item 
		group by diet_id, nutrient
), 
diet_nutrients as (
	select diet_id, 
		sum(case when nutrient = 'carbs' then 4.1 * quantity end) carb_calories, 
		sum(case when nutrient = 'energy' then quantity end) calories 
		from nut group by diet_id
)
select diet_id, (100*carb_calories/calories) as pct_calories_carbs from diet_nutrients;"

sqldf(sql)
```
