---
title: "EMR Text Field Simulation"
author: "Bob Horton"
date: "Wednesday, April 13, 2016"
output: html_document
---

We want to use the information we know about each patient to engineer an "Admitting History" (and eventually a "Discharge Note") for each patient. The goal is to develop materials usable for natural language processing demos.

```{r setup, echo=FALSE}
DATA_DIR <- "781670.f1"
diabetes_data <- read.csv(file.path(DATA_DIR, "diabetic_data_initial.csv"), na.strings = c("?", "None"))

# apply metadata from id_mapping.csv
# make sure all integer levels are covered, so we can just factor the discriptions in order
category_values <- list(
admission_source_id = read.csv(text="admission_source_id,description
1,Physician Referral
2,Clinic Referral
3,HMO Referral
4,Transfer from a hospital
5,Transfer from a Skilled Nursing Facility (SNF)
6,Transfer from another health care facility
7,Emergency Room
8,Court/Law Enforcement
9,Not Available
10,Transfer from critial access hospital
11,Normal Delivery
12,Premature Delivery
13,Sick Baby
14,Extramural Birth
15,Not Available
16,NULL
17,NULL
18,Transfer From Another Home Health Agency
19,Readmission to Same Home Health Agency
20,Not Mapped
21,Unknown/Invalid
22,Transfer from hospital inpt/same fac reslt in a sep claim
23,Born inside this hospital
24,Born outside this hospital
25,Transfer from Ambulatory Surgery Center
26,Transfer from Hospice"),

discharge_disposition_id = read.csv(text="
discharge_disposition_id,description
1,Discharged to home
2,Discharged/transferred to another short term hospital
3,Discharged/transferred to SNF
4,Discharged/transferred to ICF
5,Discharged/transferred to another type of inpatient care institution
6,Discharged/transferred to home with home health service
7,Left AMA
8,Discharged/transferred to home under care of Home IV provider
9,Admitted as an inpatient to this hospital
10,Neonate discharged to another hospital for neonatal aftercare
11,Expired
12,Still patient or expected to return for outpatient services
13,Hospice / home
14,Hospice / medical facility
15,Discharged/transferred within this institution to Medicare approved swing bed
16,Discharged/transferred/referred another institution for outpatient services
17,Discharged/transferred/referred to this institution for outpatient services
18,NULL
19,Expired at home. Medicaid only, hospice.
20,Expired in a medical facility. Medicaid only, hospice.
21,Expired, place unknown. Medicaid only, hospice.
22,Discharged/transferred to another rehab fac including rehab units of a hospital .
23,Discharged/transferred to a long term care hospital.
24,Discharged/transferred to a nursing facility certified under Medicaid but not certified under Medicare.
25,Not Mapped
26,Unknown/Invalid
27,Discharged/transferred to a federal health care facility
28,Discharged/transferred/referred to a psychiatric hospital of psychiatric distinct part unit of a hospital
29,Discharged/transferred to a Critical Access Hospital (CAH).
30,Discharged/transferred to another Type of Health Care Institution not Defined Elsewhere
"),

admission_type_id=read.csv(text="admission_type_id,description
1,Emergency
2,Urgent
3,Elective
4,Newborn
5,Not Available
6,NULL
7,Trauma Center
8,Not Mapped")
)

for (cv_name in names(category_values)){
  cv <- category_values[[cv_name]]
  diabetes_data[[cv_name]] <- factor(cv[diabetes_data[[cv_name]],"description"])
}

```

```{r load_bmi_and_sensor_assignments_data}

SENSOR_ASSIGNMENTS_FILE <- "sim_sensor_assignments.csv"
sensor_assignments <- read.csv(SENSOR_ASSIGNMENTS_FILE, 
                               colClasses=c("integer", "character", "factor", "factor"))

BMI_FILE <- "sim_bmi_data.csv"
bmi_data <- read.csv(BMI_FILE)

# bmi_data$readmitted <- diabetes_data[diabetes_data$patient_nbr==bmi_data$patient_nbr, "readmitted"]
# with(bmi_data, plot(bmi ~ readmitted))
```

# Medications

Medication columns have values "Down", "Steady", "Up", or "No", depending on how the dose has changed (if the patient takes this drug).

```{r medications}
medication_columns <- names(diabetes_data)[25:47]
lapply(medication_columns, function(col) levels(diabetes_data[[col]]))

drug_matrix <- as.matrix(diabetes_data[medication_columns]) != "No"
num_diabetes_medications <- rowSums(drug_matrix)

# plot(factor(num_diabetes_medications), diabetes_data$num_medications)

```


# Simulation of 'History of Present Illness' and 'Discharge Summary'

Also give patients fake names. This has already been run: patient names are in `patient_table.csv`.

```{r history_present_illness, eval=FALSE}
guess_age <- function(age_cat, seed=123){
  set.seed(seed)
  library(dplyr)
  age_cat %>% as.character %>% 
    (function(label) substr(label, 2, nchar(label) - 1)) %>%
    strsplit('-') %>% lapply(as.integer) %>% 
    lapply(function(v) sample(seq(v[1], v[2]-1), 1)) %>%
    unlist
}

# # Deprecated; use "pick_baby_names"
# pick_baby_name <- function(sex, yob, seed=123){
#   # to do: vectorize to get a set of names for a given sex and age
#   set.seed(seed)
#   library(babynames)
#   name_rnd <- runif(1)
#   if (yob > max(babynames$year)) yob <- max(babynames$year)
#   candidates <- babynames[babynames$year==yob & babynames$sex == sex,]
#   cum_frac <- cumsum(candidates$n)/sum(candidates$n)
#   candidates[1 + sum(cum_frac < name_rnd), ]$name
# }

pick_baby_names <- function(sex_vec, yob_vec, seed=123){
  # vectorized: gets a set of names given sex and age vectors
  sex_vec <- factor(sex_vec)
  set.seed(seed)
  library(babynames)
  yob_vec[yob_vec > max(babynames$year)] <- max(babynames$year)
  yob_vec[yob_vec < min(babynames$year)] <- min(babynames$year)
  name_vec <- character(max(length(sex_vec), length(yob_vec)))
  for (sex in levels(sex_vec)){
    for (yob in unique(yob_vec)){
      positions <- sex_vec==sex & yob_vec==yob
      N <- sum(positions)
      candidates <- babynames[babynames$year==yob & babynames$sex == sex,]
      choices <- sample(candidates$name, N, replace=TRUE, prob=candidates$n)
      name_vec[positions] <- choices
    }
  }
  name_vec
}

# # Deprecated; use "get_surnames"
# get_surname <- function(N, seed=123){
#   # to do: vectorize to get a set of surnames (for a given race)
#   set.seed(seed)
#   surnames <- read.delim("Hx/us_surnames.tsv", stringsAsFactors=FALSE)
#   # from https://en.wikipedia.org/wiki/List_of_most_common_surnames_in_North_America#United_States_.28American.29
#   surnames <- surnames[order(surnames$Rank),]
#   surnames <- transform(surnames, count=as.integer(gsub(',', '', as.character(Total.number))))
#   pop_size <- sum(surnames$count)
#   surnames$pop_fraction <- surnames$count/pop_size
#   surnames$cum_fraction <- cumsum(surnames$pop_fraction)
#   surname_rnd <- runif(N)
#   surname_idx <- sapply(surname_rnd, function(sr) 1 + sum(surnames$cum_fraction < sr))
#   surnames$Name[surname_idx]
# }

get_surnames <- function(race_vec, seed=123){
  # vectorized to get a set of surnames (for a given race)
  set.seed(seed)
  race_vec[is.na(race_vec)] <- "Other"
  surnames <- read.delim("Hx/us_surnames.tsv", stringsAsFactors=FALSE)
  # from https://en.wikipedia.org/wiki/List_of_most_common_surnames_in_North_America#United_States_.28American.29
  names(surnames)[8:13] <- c("Caucasian", "AfricanAmerican", "Asian", "NativeAmerican", "Other", "Hispanic")
  surname_vec <- character(length(race_vec))
  
  for (race in race_vec){
    positions <- race_vec == race
    N <- sum(positions)
    surname_vec[positions] <- with(surnames, sample(Name, N, replace=TRUE, prob=surnames[[race]]))
  }
  
  surname_vec
}

# diabetes_data[c("weight", "bmi")] <- bmi_data[bmi_data$patient_nbr==diabetes_data$patient_nbr, c("weight", "bmi")]
# diabetes_data$age <- guess_age(diabetes_data$age)
# 
# # diabetes_data$first_name <- mapply(pick_baby_name, ifelse(diabetes_data$gender=="Female", "F", "M"), 2016 - diabetes_data$age)
# rxOptions(numCoresToUse=4)
# rxSetComputeContext("localpar")
# first_name <- rxExec(pick_baby_name, rxElemArg(ifelse(diabetes_data$gender=="Female", "F", "M")), rxElemArg(2016 - diabetes_data$age))
# 
# diabetes_data$first_name <- unlist(first_name)
# diabetes_data$last_name <- get_surname(nrow(diabetes_data))
# 
# saveRDS(diabetes_data, file="decorated_diabetes_data.rds")

patient_surnames <- get_surnames(diabetes_data$race); saveRDS(patient_surnames, "patient_surnames.Rds")
sex_vec <- ifelse(diabetes_data$gender=="Female", "F", "M")
age_vec <- guess_age(diabetes_data$age)
# yob_vec <- 2016 - age_vec
# now <- as.Date(Sys.time())

discharge_date <- as.Date(sensor_assignments$discharge_time)

birthdates <- discharge_date - 365.24*age_vec - sample(0:364, length(age_vec), replace=TRUE)
age2 <- floor(as.integer(discharge_date - birthdates)/365.24)
hist(age2 - age_vec) # very rarely off by one

yob <- format(birthdates, format="%Y")
patient_firstnames <- pick_baby_names(sex_vec, yob)

patient_personal <- data.frame(
  patient_nbr = diabetes_data$patient_nbr,
  firstname=patient_firstnames,
  lastname=patient_surnames,
  dob = birthdates
)

# We just made one person record per encounter. But since patients can have multiple encounters, we should have only one person record per patient_nbr.
patient_table <- patient_personal[!duplicated(patient_personal$patient_nbr),]

write.csv(patient_table, file="patient_table.csv", row.names=FALSE)

```


```{r}
patient_table <- read.csv(file="patient_table.csv", stringsAsFactors=FALSE)
rownames(patient_table) <- patient_table$patient_nbr

# http://stackoverflow.com/questions/3611314/calculating-ages-in-r, by Jim
age = function(from, to) {
  from_lt = as.POSIXlt(from)
  to_lt = as.POSIXlt(to)

  age = to_lt$year - from_lt$year

  ifelse(to_lt$mon < from_lt$mon |
         (to_lt$mon == from_lt$mon & to_lt$mday < from_lt$mday),
         age - 1, age)
}

decorate_diabetes_data <- function(d_data, pt_tbl, bmi_dat, sensor_assign){
  # add first_name, last_name, weight, bmi, and numeric age
  d_data[c("height", "weight", "bmi")] <- bmi_dat[bmi_dat$patient_nbr==d_data$patient_nbr, 
                                                   c("height", "weight", "bmi")]
  d_data[c("first_name", "last_name", "dob")] <- pt_tbl[as.character(d_data$patient_nbr), 
                                                        c("firstname", "lastname", "dob")]
  
  d_data$discharge_time <- sensor_assign[sensor_assign$encounter_id==d_data$encounter_id, "discharge_time"]
  
  d_data$age <- with(d_data, {
    discharge <- as.Date(discharge_time) # %>% ymd
    dob <- dob # %>% ymd
    age(dob, discharge)
  })
  
  d_data
}
    # floor((as.Date(discharge_time) - as.Date(dob))/365.24)
    # library(lubridate)
    # library(dplyr)
    # discharge <- as.Date(discharge_time) %>% ymd
    # dob <- dob %>% ymd
    # age(dob, discharge)
    ## Bug in lubridate: interval 1964-02-29 UTC--2016-02-27 UTC breaks period(., unit="year")
    ## my_interval <- as.interval(ymd("2016-02-27") - ymd("1964-02-29"), start=ymd("1964-02-29"))
    ## as.period(my_interval, unit="month") # this works
    ## as.period(my_interval, unit="year") # error
    
    #     as.interval(discharge - dob, dob) %>% 
    #       as.period(unit="year") %>% 
    #       as.character %>% 
    #       gsub("y.*$", "", .)
    
    # unit="month",  %>% "/"(12) fails on different cases, 

decorated_diabetes_data <- decorate_diabetes_data(diabetes_data, patient_table, bmi_data, sensor_assignments)

```


# Admitting History and Physical

See more examples (here)[https://www.med.unc.edu/medselect/resources/sample-notes].

See this example of a (discharge summary)[http://clerkship.medicine.ufl.edu/portfolio/interpersonal-and-communicative-skills/discharge-summarytransfer-noteoff-service-note-instructions/].

```{r}
# encounter_record <- diabetes_data[which.max(num_diabetes_medications),]

sim_history_present_illness <- function(dd_data){
  with(dd_data, {
    
      nominative_pronoun <- ifelse(gender=="Female", "she", "he")
      posessive_pronoun <- ifelse(gender=="Female", "her", "his")
      sex <- ifelse(gender=="Female", ifelse(age < 18, "girl", "woman"), ifelse(age < 18, "boy","man"))
      
      er_use <- cut(number_emergency, breaks=c(0, 1, 10, 20, 30, Inf), 
                    labels=c("has never been", "is occasionally", "is frequently", 
                             "is very often", "is constantly being"), include_lowest=TRUE)
      
      get_er_variant <- function(N) 
                          sample(c("emergency room", "ER", "emergency department", "ED"), N, replace=TRUE)
      er_variant <- get_er_variant(length(er_use))
      er_use_clause <- paste(nominative_pronoun, er_use, er_variant, ".")

      weight_category <- cut(bmi, breaks=c(0, 18.5, 25, 30, 35, 40, 50, Inf), 
                             labels=c("underweight", "normal weight", "overweight", 
                                      "mildly obese", "moderately obese", 
                                      "severely obese", "super obese"), include.lowest=TRUE)

      basis_clause <- ifelse(admission_type_id %in% c("Elective", "Emergency", "Urgent"), 
             paste("on an", tolower(admission_type_id), "basis"), "")
      
      er_variant2 <- get_er_variant(length(er_use))
      
      admission_source_clause <- ifelse(admission_source_id %in% 
                                          c("Clinic Referral", "Transfer from another health care facility", 
                                            "Transfer from a hospital", "Physician Referral"), 
                                        paste("as a", tolower(admission_source_id), basis_clause), 
                                        ifelse(admission_source_id == "Emergency Room", paste("from the", er_variant2), ''))

      levels(medical_specialty) <- gsub("([a-z])([A-Z])", "\\1 \\2", 
                                        gsub("([a-z])(with|and)","\\1 \\2", 
                                             gsub("-", " ", levels(medical_specialty))), perl=TRUE)
      medical_specialty_clause <- paste("admitted to", 
                                        ifelse(is.na(medical_specialty), 
                                               'the hospital', 
                                               paste("the", tolower(medical_specialty), "service")))
      
      hpi1 <- paste("This is a", age, "year old", weight_category, race, sex, medical_specialty_clause, admission_source_clause, ".")
      
      history_present_illness <- gsub("\\s+", " ", gsub(" \\.", ".", hpi1))
      
      history_present_illness
  })
}

sim_medications <- function(dd_data){
  # Count individual components of combination meds
  medication_columns <- names(diabetes_data)[25:47]
  get_medications <- function(encounter_table){
    lapply(1:nrow(encounter_table), function(i){
        meds <- medication_columns[encounter_table[i, medication_columns] != "No"]
        meds <- unique(unlist(strsplit(meds, "\\.")))
        meds
      }
    )
  }
  medications <- sapply(get_medications(dd_data), function(meds) paste(meds, collapse="\n\t*"))
  sapply(medications, function(meds) if (""==meds) "NONE" else paste0("\n\t*", meds))
}

sim_assessment <- function(dd_data){
  ICD9_DIR <- "ICD-9-CM-v32-master-descriptions"
  icd9_dx <- read.csv(file.path(ICD9_DIR, "CMS32_DESC_LONG_SHORT_DX.csv"), 
                      colClasses=rep("character", 4)) #stringsAsFactors=FALSE
  
  # ICD9 codes are represented without decimal points in the lookup table. 
  # Leading zeros are significant, but they are missing in our dataset. Some trailing zeros are missing as well.
  # Eg, (04110 and 4110), (01570 and 1570) are different codes. So I have to guess in some cases.
  lookup_icd9_dx <- function(icd9){
    codes <- levels(icd9)[icd9]
    nums <- grepl("\\.", codes)  # !grepl("^[EV]", codes) & !is.na(codes)
    codes[nums] <- as.character(as.numeric(codes[nums]) * 100)
    names(codes) <- codes
    key <- gsub("^0+", "", icd9_dx$`DIAGNOSIS.CODE`)
    vsubs <- grep("V.+00$", icd9_dx$DIAGNOSIS.CODE, value=TRUE)
    names(vsubs) <- gsub("00$","",vsubs)
    substitutions <- c(E909="E9090", '411'="4110", vsubs)
    lapply(names(substitutions), function(n) codes[codes==n] <- substitutions[n])
    res <- sapply(codes, function(code) icd9_dx$LONG.DESCRIPTION[which(key==code)])
    # guess again by appending one 0 to the codes with missing values
    missing <- sapply(res, length)==0
    codes[missing] <- paste0(codes[missing], "0")
    res <- sapply(codes, function(code) icd9_dx$LONG.DESCRIPTION[which(key==code)])
    # guess yet again on missing values by appending yet another 0
    missing <- sapply(res, length)==0
    codes[missing] <- paste0(codes[missing], "0")
    sapply(codes, function(code){ 
      v <- icd9_dx$LONG.DESCRIPTION[which(key==code)]
      v[length(v)]  # if there are multiple matches, return the last one (because 411 maps to 4110 not 04110)
    })
  }
  
  icd9_matrix <- cbind (
    lookup_icd9_dx(dd_data$diag_1),
    lookup_icd9_dx(dd_data$diag_2),
    lookup_icd9_dx(dd_data$diag_3)
  )
  
  diagnoses <- apply( icd9_matrix, 1, paste, collapse="\n\r*" )
  
  with(dd_data,
    ifelse (number_diagnoses > 3, paste(diagnoses, sprintf("\n\r* and %d others", number_diagnoses - 3)), diagnoses)
  )
}

sim_physical_exam <- function(dd_data){
  N <- nrow(dd_data)
  #VITAL SIGNS: Weight 175.2 pounds which is 79.3 kg, blood pressure 142/96 by the nurse, 140/92 by me, pulse is 64.
  template <- "Vital signs: blood pressure %d/%d, pulse %d, respiratory rate %d, weight %0.1f %s, height %s, BMI %d."
  bp_systolic <- round(rnorm(N, mean=120, sd=15))
  bp_diastolic <- round(rnorm(N, mean=75, sd=10))
  pulse <- round(rnorm(N, mean=80, sd=10))
  respirarory_rate <- round(rnorm(N, mean=18, sd=3))
  weight_units <- sample(c("kg", "kilograms", "lb", "pounds"), N, replace=TRUE)
  with(dd_data,{
    weight[is.infinite(weight)] <- NA
    bmi[is.infinite(bmi)] <- NA
    reported_weight <- ifelse (weight_units %in% c("lb", "pounds"), weight * 2.2, weight)
    # height <- sqrt(weight/bmi) # in meters
    reported_height <- sprintf("%dft %din", floor(39.3701 * height/12), round((39.3701 * height) %% 12))
    sprintf(template, bp_systolic, bp_diastolic, pulse, respirarory_rate, reported_weight, weight_units, reported_height, round(bmi))
  })
}

sim_admitting_history <- function(dd_data){
  template <- "## History of present illness\n%s

## Physical Exam\n%s

## Medications\n%s

## Assessment\n%s

"
  sprintf( template, 
    sim_history_present_illness(dd_data),
    sim_physical_exam(dd_data),
    sim_medications(dd_data),
    sim_assessment(dd_data)
  )
}

sim_discharge_summary <- function(dd_data){
  with(dd_data,{
    
    discharge_disposition <- as.character(discharge_disposition_id)
    
    gender_title <- ifelse(gender=="Female", "Ms.", "Mr.")
    this_hospital_phrase <- sample(c("in this facility", "in the hospital", "under our care"), 
                                   nrow(dd_data), replace=TRUE)
    substring(discharge_disposition, 1) <- tolower(substring(discharge_disposition, 1))
    joiner <- ifelse(grep("^(admitted|discharged)", discharge_disposition), "was", 
              ifelse(grep("hospice", discharge_disposition), "was transfered to", ""))
   
    sprintf("After %d days %s, %s %s %s %s.", 
            time_in_hospital, this_hospital_phrase, 
            gender_title, last_name, joiner, discharge_disposition)
  })
}

## Discharge Summary\n%s
# if(length(ds <- sim_discharge_summary(dd_data)) == 0) "None" else ds

```

```{r debug_sim_admitting_history}
# try processing blocks; some will fail, but this is a fast way to do the rest
N <- nrow(decorated_diabetes_data)
hx <- character(N)
block_size <- 1000
for (k in 1:floor(N/block_size)){
  i <- block_size * (k-1) + 1
  rng <- i:(i+(block_size-1))
  print(sprintf("block from %d to %d\n", rng[1], rng[block_size]))
  hx[rng] <- tryCatch(sim_admitting_history(decorated_diabetes_data[rng,]), error=function(e) NA)
}
rng <- (i + block_size):N
print(sprintf("last few from %d to %d\n", min(rng), max(rng)))
hx[rng] <- tryCatch(sim_admitting_history(decorated_diabetes_data[rng,]), error=function(e) NA)

# plot(is.na(hx), pch='.')
sum(is.na(hx))

for (i in which(is.na(hx))){
  tryCatch({
      print(sprintf("i = %d\n", i))
      if (is.na(hx[i])) hx[i] <- sim_admitting_history(decorated_diabetes_data[i,])
    },
    error = function(e){NA}
  )
}

# plot(is.na(hx), pch='.')
sum(is.na(hx)) # 14
redo <- which(is.na(hx))


decorated_diabetes_data[redo,]
# All have infinite weight (and bmi).
# BUT not all records with infinite weight ended up on redo list.  (1019 total)

md2html <- function(markdown_vec){
  library(markdown)
  library(dplyr)
  html_vec <- sapply(markdown_vec, function(txt){
    if (is.na(txt) || length(txt) == 0 || txt == ""){
      ""
    } else {
        renderMarkdown(text=txt) %>% 
              gsub("\\n", "<br />", .) %>%
              gsub("\\t", "    ", .)
    }
  })
  unname(html_vec)
}

hx_html <- md2html(hx)
admitting_history <- data.frame(encounter_id = decorated_diabetes_data$encounter_id, hx=hx_html, stringsAsFactors=FALSE)
write.table(admitting_history, "admitting_history.csv", sep=",", row.names=FALSE, col.names=FALSE)

# admitting_history <- read.table("admitting_history.csv", sep=",", col.names=c("encounter_id", "hx"), stringsAsFactors=FALSE)

sample_history <- rbind(admitting_history[1:100,], admitting_history[redo[1],])
# 100 normal, and 1 NA
write.table(sample_history, "sample_history.csv", sep=",", row.names=FALSE, col.names=TRUE)

```


## Extract weight in kilograms from text

```{r extract_weight}
extract_weight_in_kg <- function(txt){
  pat <- "Vital signs:[^\\n<]+weight ?([0-9\\.]+) ?(kg|kilo|kilogram|lb|pound)"
  matches <- regexec(pat, txt, ignore.case=TRUE)
  
  get_matches <- function(match, src)
    substring(src, first=match, last=match + attr(match, "match.length"))
  
  M_list <- mapply(get_matches, matches, txt, SIMPLIFY=FALSE) # I'll simplify it myself, thank you
  M <- do.call("cbind", lapply(M_list, function(v) v[1:3]))
  
  val <- as.numeric(M[2,])
  unit <- ifelse(substr(M[3,],1,1) == 'k', 1, 1/2.20462)
  
  val * unit
}

wt_in <- decorated_diabetes_data$weight
wt_out <- extract_weight_in_kg(hx) # takes about 6 seconds
plot(wt_in, wt_out, pch='.')
hist(wt_in/wt_out, breaks=100)
sum(is.na(wt_in)) #3
sum(is.na(wt_out)) # 1037

# hx[vanished[1:1]] "weight Inf kilograms"
```

# To Do for Next Iteration

## Adding Symptoms

I have not found a perfetly suitable open source database of symptoms to accompany ICD9 diagnoses, but the body system indicators from the (HCUPS Chronic Condition Indicator)[https://www.hcup-us.ahrq.gov] database may be helpful:

```{r symptoms}
body_system_indicator_category <- c(
  '1' = 'Infectious and parasitic disease',
	'2' = 'Neoplasms',
	'3' = 'Endocrine, nutritional, and metabolic diseases and immunity disorders',
	'4' = 'Diseases of blood and blood-forming organs',
	'5' = 'Mental disorders',
	'6' = 'Diseases of the nervous system and sense organs',
	'7' = 'Diseases of the circulatory system',
	'8' = 'Diseases of the respiratory system',
	'9' = 'Diseases of the digestive system',
	'10' = 'Diseases of the genitourinary system',
	'11' = 'Complications of pregnancy, childbirth, and the puerperium',
	'12' = 'Diseases of the skin and subcutaneous tissue',
	'13' = 'Diseases of the musculoskeletal system',
	'14' = 'Congenital anomalies',
	'15' = 'Certain conditions originating in the perinatal period',
	'16' = 'Symptoms, signs, and ill-defined conditions',
	'17' = 'Injury and poisoning',
	'18' = 'Factors influencing health status and contact with health services'
)

cci <- read.csv("Hx/cci2015.csv", skip=1)
# category descriptions: chronic (1) versus not chronic (0)


```

### HCUPS ICD9 Keys

* Alphanumeric diagnosis codes must be left-justified.
* Two spaces must always follow a 3-character diagnosis code.
* One space must always follow a 4-character diagnosis code.
* Trailing blanks should never be zero-padded (i.e., blank spaces following a 3-character diagnosis code should not be filled with zeroes).
* Leading zeroes must be preserved; they are significant.