---
title: "Simulating wearable sensor data"
author: "Bob Horton"
date: "Wednesday, March 23, 2016"
output: html_document
---

Using the multivariate normal model developed for another project, I will simulate sensor readings to complement the EMR data from the diabetes hospital readmission study.


```{r}
NUM_PATIENTS <- 100
NUM_STATES <- 30 * 24 * 60 # one state per minute
PATIENT_STATE <- matrix(numeric(NUM_PATIENTS * NUM_STATES), nrow=NUM_PATIENTS)

```


Considering the postures as states, I will make a set of state-transition matrixes. These will depend on how sedentary an individual is, as well as time of day. There will be different matrixes for 

```{r posture_transition_model}

reported_postures <- c(
   horizontal = 0,
   upright = 2,
	 walking = 3,
	 running = 4,
	 unknown = 5
)

simulated_postures <- c("sleeping", "resting", "sitting", "standing", "walking", "running")

posture_mapping <- c(
  sleeping = 'horizontal',
  resting = 'horizontal',
  sitting = 'upright',
  standing = 'upright',
  walking = 'walking',
  running = 'running'
)

# I won't model "unknown"; postures near transitions will have some probability of being reported "unknown".

active_night <- matrix(c(
      0.95, 0.05, 0.00, 0.00, 0.00, 0.00, # sleeping
      0.45, 0.45, 0.10, 0.00, 0.00, 0.00, # resting
      0.00, 0.30, 0.55, 0.15, 0.00, 0.00, # sitting
      0.00, 0.00, 0.55, 0.10, 0.45, 0.00, # standing
      0.00, 0.00, 0.00, 0.75, 0.20, 0.05, # walking
      0.00, 0.00, 0.00, 0.00, 0.05, 0.95  # running
    ), nrow=6, byrow=T)

active_day <- matrix(c(
      0.25, 0.75, 0.00, 0.00, 0.00, 0.00, # sleeping
      0.05, 0.45, 0.50, 0.00, 0.00, 0.00, # resting
      0.00, 0.10, 0.55, 0.35, 0.00, 0.00, # sitting
      0.00, 0.00, 0.25, 0.60, 0.15, 0.00, # standing
      0.00, 0.00, 0.00, 0.35, 0.50, 0.15, # walking
      0.00, 0.00, 0.00, 0.00, 0.10, 0.90  # running
    ), nrow=6, byrow=T)

sedentary_night <- matrix(c(
      0.95, 0.05, 0.00, 0.00, 0.00, 0.00, # sleeping
      0.25, 0.45, 0.30, 0.00, 0.00, 0.00, # resting
      0.00, 0.20, 0.75, 0.05, 0.00, 0.00, # sitting
      0.00, 0.00, 0.75, 0.10, 0.25, 0.00, # standing
      0.00, 0.00, 0.00, 0.75, 0.20, 0.05, # walking
      0.00, 0.00, 0.00, 0.00, 0.95, 0.05  # running
    ), nrow=6, byrow=T)

sedentary_day <- matrix(c(
      0.25, 0.75, 0.00, 0.00, 0.00, 0.00, # sleeping
      0.25, 0.45, 0.30, 0.00, 0.00, 0.00, # resting
      0.00, 0.20, 0.65, 0.15, 0.00, 0.00, # sitting
      0.00, 0.00, 0.35, 0.40, 0.15, 0.00, # standing
      0.00, 0.00, 0.00, 0.55, 0.40, 0.05, # walking
      0.00, 0.00, 0.00, 0.00, 0.80, 0.20  # running
    ), nrow=6, byrow=T)

POSTURE_TRANSITION_MATRICES <- list(
  active = list(
    day = active_day,
    night=active_night
  ),
  sedentary = list(
    day = sedentary_day,
    night=sedentary_night
  )
)

next_state <- function(current_state, transition_matrix)
  which.max(transition_matrix[current_state,] * runif(nrow(transition_matrix)))

```


```{r run_state_simulation}
# Run simulation of states

# The year starts at midnight with everyone asleep
PATIENT_STATE[,1] <- rep(1, NUM_PATIENTS)

#
PATIENT_LIFESTYLE <- sample(c("active", "sedentary"), NUM_PATIENTS, replace=TRUE, prob=c(0.45, 0.55))
PATIENT_CIRCADIAN_ADJUSTMENT <- round(rnorm(NUM_PATIENTS, mean=0, sd=45))

SIM_START_TIME <- as.POSIXct("2016-01-01 00:00:00")
TIMELINE <- SIM_START_TIME + (1:NUM_STATES) * 60 
DAY_TIMES <- TIMELINE[1:(24*60)]
DAY_PART <- ifelse(as.POSIXlt(DAY_TIMES)[['hour']] %in% 7:22, "day", "night")

sim_state <- function(){
  for (pt in 1:NUM_PATIENTS){
    pt_tmats <- POSTURE_TRANSITION_MATRICES[[PATIENT_LIFESTYLE[pt]]]
    for (st in 2:NUM_STATES){
      shifted_state_idx <- 1 + (st + (PATIENT_CIRCADIAN_ADJUSTMENT[pt] - 1)) %% length(DAY_TIMES)
      circadian_day_part <- DAY_PART[shifted_state_idx]
      transmat <- pt_tmats[[circadian_day_part]]
      PATIENT_STATE[pt, st] <<- next_state(PATIENT_STATE[pt, st-1], transmat)
    }
  }
}


```


Each patient has a circadian shift, normally distributed with mean 0 and standard deviation 45 minutes. May be easiest to start out everyone the same for a day 

To Do:

Add a mealtime daypart? Also an eating state? It could map to sitting for posture reporting. Other dayparts could also have eating, which would represent snacks. Blood sugar will be a lagging window function over eating states. Meal constituents will vary by individual, and affect blood sugar.
